Type Error Telemetry at Scale

### 2023-09-28 programming23/paper.tex
sec 3

Telemetry allows an application to ``phone home'' with data
summarizing usage patterns, such performance data, crash reporting,
and feature uptake. A typical usage is in deciding whether an API can
be deprecated: without telemetry it may be difficult to know how
popular a feature is with users, but with telemetry it is
straightforward.

IDEs use telemetry in a similar fashion to most user-facing
applications, for example VSCode~\cite{vsc-telemetry} and
IntelliJ~\cite{intellij-telemetry} report telemetry.
See~\cref{s:related} for futher comparisons among telemetry designs.

Telemetry for programming languages can be controversial, for example
the lively discussion around telemetry in the Go
toolchain~\cite{golang-telemetry}. 
The Luau open source toolchain does \emph{not} report telemetry,
as it is designed to be used in build environments such as Continuous Integration
servers, where hermetic deterministic builds are expected.

Much of the concern about telemetry is because of its potential impact on
privacy, by revealing Personally Identifiable Information (PII).
For this reason, it is important that any experiments using telemetry
not reveal PII:
no source code;
no source code locations;
no error messages (which may contain source code);
no record of the creator's identity, locale, or IP address;
and no information about what creation the data came from.
A good introduction to the privacy implications and tradeoffs
involved with telemetry is~\cite{transparent-telemetry}.

For performance reasons, there are limits to telemetry data:
telemetry records have size limits, and the performance impact of
recording telemetry should be minimal.

In summary, the limitations of telemetry are:
\begin{itemize}
  \item Data which reveals no PII.
  \item Fixed-size telemetry records.
  \item Minimal overhead to record telemtry.
\end{itemize}
In addition, due to the architecture of Roblox Studio
there is some information that is not available to the
typechecker:
\begin{itemize}
  \item The typechecker does not have access to some lifecycle events,
    such as save, quit and publish.
  \item The typechecker does not have access to GUI state, for example
    whether the Script Analysis widget is visible.
\end{itemize}

@@

\begin{figure}[t]\centering
  \begin{tabular}{|l|l|l|l|l|l|l|}\hline
    ID & Time & Mode & Reason & Size Info & Overall Counts & Edit Range Counts
    \\\toprule
    \multicolumn{6}{c}{\upbracefill} & \multicolumn{1}{c}{\upbracefill} \\[-0.5ex]
    \multicolumn{6}{c}{fixed-size} & \multicolumn{1}{c}{variable-size}
  \end{tabular}
  \caption{Structure of a telemetry record. There are 13 overall counts and up to 70 edit range counts in each record.}
  \label{f:tdata}
\end{figure}

The telemetry for a Roblox Studio session is gathered as a series of
telemtry records, all with the same \panon{} session
identifier. This allows us to correlate telemetry across a single
session, but not between sessions.

In order to avoid swamping our telemetry servers, we randomly throttle
the generation of telemetry records:
\begin{itemize}
  \item
    1\% of Roblox Studio sessions were enrolled in generating telemetry, and
  \item
    0.5\% of runs of the typechecker (approximately 1 in 200 keystrokes)
      in an enrolled session generated a telemetry record.
\end{itemize}
In order to find how file-switching correlated with type errors, we additionally have:
\begin{itemize}
  \item
    every run of the typechecker in an enrolled session which switched file generated a telemetry record.
\end{itemize}

@@

Each telemetry record contains seven kinds of data.
Refer to~\cref{f:tdata} for an illustration:
\begin{enumerate}
  \item
    ID: a \panon{} identifier for the current session
  \item
    Time: one timestamp from the client and one from the server.
    %% timestamp explanation \cref{s:data-cleaning}
  \item
    Mode: type analysis mode of the current file, which is either \mnocheck{},
    \mnonstrict{}, or \mstrict{}.
  \item
    Reason: a flag that explains whether this record was sent due to a
    randomly-selected keystroke or a module switch.
  \item
    Size Info: number of lines in the codebase and number of lines in the edit range.
  \item
    Overall Counts: summarizes type errors and \FS{} errors during the
    last two invocations of type analysis, and additionally reports the number
    of times type analysis hit an internal limit on problem
    size~(see~\cref{s:code-too-complex} for details).
    For each invocation and each kind of
    analysis error, there are three summary counts:
    %% grand total = 12
    \begin{itemize}
      \item total number of errors,
      \item errors in the current module, and
      \item errors in the current edit range.
    \end{itemize}
  \item
    Edit Range Counts: these list specific errors that arose in the last two
    type analyses and overlap with the current edit range.
    Since there are 35 possible errors~(\cref{s:context}), there can be up to 70 counts
    in a telemetry record.
    Exactly how to interpret this data depends on which analysis it came from:
    \begin{itemize}
      \item
        Errors in the latest type analysis are currently visible to the
        creator.
        These may have been introduced by the changes in the edit range.
      \item
        Errors from the previous type analysis were visible before the creator
        made the latest edits.
        These may have motivated some of the changes in the edit range.
    \end{itemize}
\end{enumerate}

>>>>

Telemetry allows an application to ``phone home'' with data
summarizing usage patterns, such performance data, crash reporting,
and feature uptake. A typical usage is in deciding whether an API can
be deprecated: without telemetry it may be difficult to know how
popular it is, but with telemetry it is
straightforward.

IDEs use telemetry in a similar fashion\todo{I really don't know what
  this means. Are you saying VSCode is constantly calling home? Is it
  telling Racket about my API usage? Do we need anything in this para?} to most user-facing
applications, for example VSCode~\cite{vsc-telemetry} and
IntelliJ~\cite{intellij-telemetry} report telemetry.
See~\cref{s:related} for futher comparisons among telemetry designs.

Telemetry for programming languages can be controversial:e.g., see
the lively discussion around telemetry in the Go
toolchain~\cite{golang-telemetry}. 
The Luau open source toolchain does \emph{not} report telemetry,
as it is designed to be used in build environments such as Continuous Integration
servers, where hermetic deterministic builds are expected.

Telemetry therefore faces at least two major constraints:
\begin{itemize}

\item It should not reveal private information. This can include
  Personally Identifying
  Information (PII) about creators: their identity, location, etc. It also
  includes trade secrets. Even an error message that contains the name
  of a function could reveal something the creator would not want to
  share.

\item For performance reasons, it should have minimal run-time
  overhead and also transmit small and bounded amounts of data.
\end{itemize}
Cox discusses in detail the privacy implications and tradeoffs
in telemetry~\cite{transparent-telemetry}.

% BEN: I believe the above subsumes all of the below:
% Much of the concern about telemetry is because of its potential impact on
% privacy. It can not only reveal Personally Identifiable Information
% (PII) but also trade secrets, etc. Thus, telemetry should not reveal
% PII, source code, error messages (which can refer to source), the
% creator's identity, and so on.
% % and no information about what creation the data came from.

% For performance reasons, there are limits to telemetry data:
% telemetry records have size limits, and the performance impact of
% recording telemetry should be minimal.

% In summary, the limitations of telemetry are:
% \begin{itemize}
%   \item Data that don't reveal private information.
%   \item Fixed-size telemetry records.
%   \item Minimal overhead to record telemtry.
% \end{itemize}

In addition, due to the architecture of Roblox Studio
there is some information that is not available to the
typechecker:
\begin{itemize}
  \item The typechecker does not have access to some lifecycle events,
    such as save, quit and publish.
  \item The typechecker does not have access to GUI state, for example
    whether the Script Analysis widget is visible.
\end{itemize}

@@

\begin{figure}[t]\centering
  % BEN: The left brace is confusing. Is ID not fixed?
  \begin{tabular}{|l|l|l|l|l|l|l|}\hline
    ID & Time & Mode & Reason & Size Info & Overall Counts & Edit Range Counts
    \\\toprule
    \multicolumn{6}{c}{\upbracefill} & \multicolumn{1}{c}{\upbracefill} \\[-0.5ex]
    \multicolumn{6}{c}{fixed-size} & \multicolumn{1}{c}{variable-size}
  \end{tabular}
  \caption{Structure of a telemetry record. There are 13 overall counts and up to 70 edit range counts in each record.}
  \label{f:tdata}
\end{figure}

The telemetry for a Roblox Studio session is gathered as a series of
records, all with the same \panon{} session
identifier. This allows us to correlate telemetry across a single
session, but not between sessions.

In order to avoid swamping our telemetry servers, we randomly throttled
the generation of telemetry records:
\begin{itemize}
  \item
    1\% of Roblox Studio sessions were enrolled in generating telemetry, and
  \item
    0.5\% of runs of the typechecker (approximately 1 in 200 keystrokes)
      in an enrolled session generated a telemetry record.
\end{itemize}
In order to find how file-switching correlated with type errors, we
additionally recorded
every run of the typechecker in an enrolled session which switched
file generated a telemetry record.\todo{I cannot parse this!}

\Cref{f:tdata} shows the structure of our telemetry records:
\begin{enumerate}
  \item
    ID: a \panon{} identifier for the current session.
  \item
    Time: one timestamp from the client and one from the server.
    %% timestamp explanation \cref{s:data-cleaning}
  \item
    Mode: type analysis mode of the current file: \mnocheck{},
    \mnonstrict{}, or \mstrict{}.
  \item
    Reason: a flag that explains whether this record was sent due to a
    randomly-selected keystroke or a module switch.
  \item
    Size Info: number of lines in the codebase and number of lines in the edit range.
  \item
    Overall Counts: summarizes type errors and \FS{} errors during the
    last two invocations of type analysis, and additionally reports the number
    of times type analysis hit an internal limit on problem
    size~(see~\cref{s:code-too-complex} for details).
    For each invocation and each kind of
    analysis error, there are three summary counts:
    %% grand total = 12
    \begin{itemize}
      \item total number of errors,
      \item errors in the current module, and
      \item errors in the current edit range.
    \end{itemize}
  \item
    Edit Range Counts: these list specific errors that arose in the last two
    type analyses and overlap with the current edit range.
    Since there are 35 possible errors~(\cref{s:context}), there can be up to 70 counts
    in a telemetry record.
    Exactly how to interpret this data depends on which analysis it came from:
    \begin{itemize}
      \item
        Errors in the latest type analysis are currently visible to the
        creator.
        These may have been introduced by the changes in the edit range.
      \item
        Errors from the previous type analysis were visible before the creator
        made the latest edits.
        These may have motivated some of the changes in the edit range.
    \end{itemize}
\end{enumerate}

@@
